---
title: "Bias - Summary of selection methods used and none sampled for;"
date: \today
output: html_document
---

```{r READ_ME, include = F}

# Change data_path and file_name in beginning of setup chunk

# Note:
# This report only handles one sampling scheme at a time
# Data need to be in the same format as the

# TODO:
# 1. Include all types of tables in the recType reference
# 2. 

```


```{r setup, include = F}

# Where to find data and name thereof
data_path <- "./WKRDB-EST2/testData/output/DBErawObj/"

file_name <- "DBErawObj_DK_1966_H1.rds"

source("./WKRDB-EST2/subGroup7/funs/doDBEestimantionObjUppMid.R")

library(dplyr)
library(knitr)
library(devtools)
install_github("ices-tools-prod/icesVocab")
library(icesVocab)
library(tidyr)
library(kableExtra)

knitr::opts_chunk$set(echo = F, message = F, warning = F)

```

```{r ref, include = F}


# Translate recType to understandable language

recType <- c("DE", "SD", "VS", "FT", "FO", "SS", "SA")

recType_name <-
  c(
    "Design",
    "Sampling details",
    "Vessel",
    "Fishing trip",
    "Fishing operation",
    "Species selection",
    "Sample"
  )

recType_ref <- data.frame(recType, recType_name)

# Ref from probabilistic and ....

selectMethod_ref <- getCodeList("RS_SelectionMethod")
selectMethod_ref <- mutate(selectMethod_ref, selectMethod_group = ifelse(Key %in% c("NPAH",
    "NPEJ",
    "NPQS"), "None_prob", 
              ifelse(Key %in% c("CENSUS",
    "SRSWOR",
    "SRSWR",
    "SYSS",
    "UPSWOR",
    "UPSWR"), "Prob", NA))) 

# RS_ReasonForNotSampling

noSampReason <- getCodeList("RS_ReasonForNotSampling")

```


```{r get_data}

h1 <- readRDS(paste0(data_path, file_name))

h1$VS$VSselectMeth[h1$VS$VSstratumName == "VS_stratum1"] <- "NPEJ"
h1$FT$FTselectMeth[h1$FT$FTid %in% sample(h1$FT$FTid, 8)] <- "NPAH"
h1$FO$FOselectMeth <- "SRSWOR"
h1$SS$SSselectMeth <- "CENSUS"
h1$SA$SAselectMeth <- "SRSWOR"

h1_test <- doDBEestimantionObjUpp(input_list = h1)
```

```{r sub_title, results="asis"}

cat("  \n#",  unique(h1_test$de$DEsampScheme), ", ", unique(h1_test$de$DEyear), " \n")
cat("  \n")
```


## Selection methods

The selection methods are grouped the following way

```{r, results = "asis"}

selectMethod_ref <-
  mutate(
    selectMethod_ref,
    selectMethod_group_name = ifelse(
      selectMethod_group == "Prob",
      "Probabilistic",
      ifelse(selectMethod_group == "None_prob", "None-probabilistic", NA)
    )
  )

kable_styling(kable(
  select(selectMethod_ref, Key, Description, selectMethod_group_name),
  col.names = c("Code", "Description", "Selection method type")
))

cat("  \n")
```

NotApp and NotSam are not considered in these overviews.

### Overall - summary

Number of samples selected in a probabilistic and none-probabilistic way per sampling level.

The use of probabilist selection is highlighted in green and number of none-probabilistic selections are highlighted in red.



```{r, echo = F}

dat <- data.frame(matrix(ncol = 4 , nrow = 0))
colnames(dat) <- c("su", "recType_name", "Prob", "None_prob")

dat$su <- as.character(dat$su)
dat$recType_name <- as.character(dat$recType_name)
dat$Prob <- as.integer(dat$Prob)
dat$None_prob <- as.integer(dat$None_prob)

for (i in c(4:length(h1_test))) {
  
  dat_0 <- left_join(h1_test[[i]], recType_ref)
  dat_0 <- left_join(dat_0, selectMethod_ref, by = c("selectMeth" = "Key"))
  dat_0 <- subset(dat_0, !selectMeth %in% c("NotApp", "NotSam"))
  dat_0$su <- gsub("su", "", dat_0$su)
  
  dat_01 <- summarise(group_by(dat_0, su, recType_name, selectMethod_group), no = length(selectMeth))
  dat_01 <- spread(dat_01, key = selectMethod_group, value = no, fill = 0)
  dat <- bind_rows(dat, dat_01)
}

heading <- distinct(left_join(h1_test$de, h1_test$sd, by = c("DEid" = "DEid")), ctry, DEyear, DEsampScheme)

summary_select <- bind_cols(heading, dat)

summary_select$None_prob[is.na(summary_select$None_prob)] <- 0
summary_select$Prob[is.na(summary_select$Prob)] <- 0

summary_select %>%
  mutate(
    Prob = cell_spec(
      Prob,
      color = ifelse(Prob > 0, "white", "white"),
      background  = ifelse(Prob > 0, "green", "red"),
      bold = ifelse(Prob > 0, T, T)
    ),
    None_prob = cell_spec(
      None_prob,
      color = ifelse(None_prob > 0, "white", "white"),
      background  = ifelse(None_prob > 0, "red", "green"),
      bold = ifelse(None_prob > 0, T, T)
    )
  )  %>%
  kable(
    booktabs = T,
    col.names = c(
      "Sampling scheme",
      "Year",
      "Country",
      "Sampling unit level",
      "Sampling unit type",
      "Number of probabilistic selections",
      "Number of none-probabilistic selections"
    ),
    escape = FALSE
  ) %>%
  kable_styling()
```

### Detailed summary per sampling level

Number per selection methods used per sampling level. 

In each table the stratum from the level above is stated to make identification easier.

#### 1. sampling unit

```{r, }

dat_1 <- left_join(left_join(h1_test$de, h1_test$sd, by = c("DEid" = "DEid")), h1_test$su1, by = c("id" = "idAbove"))

dat_1 <- left_join(dat_1, recType_ref, by = c("recType.y" = "recType"))
dat_1 <- subset(dat_1, !selectMeth %in% c("NotApp", "NotSam"))
dat_11 <- summarise(group_by(dat_1, recType_name, DEsampScheme, DEyear, ctry, DEstratumName, stratumName, selectMeth), no = length(selectMeth))


dat_11 %>%
  mutate(
    selectMeth = cell_spec(
      selectMeth,
      color = "white",
      background  = ifelse(selectMeth %in% c("NPAH", "NPEJ", "NPQS"), "red", "green"),
      bold = T
    )
  )  %>%
  kable(
    booktabs = T,
    col.names = c(
      "Sampling unit type",
      "Sampling scheme",
      "Year",
      "Country",
      "Stratum above",
      "Stratum here",
      "Selection method",
      "Number of samples"
    ),
    escape = FALSE
  ) %>%
  kable_styling()

```

```{r, results = "asis"}

for (j in c(5:length(h1_test))) {
  
  dat_2 <- left_join(h1_test[[j-1]], h1_test[[j]], by = c("id" = "idAbove"))
  
  cat("  \n#### ", gsub("su", "", unique(dat_2$su.y)), ". sampling unit \n", sep = "")
  
  dat_2 <- left_join(dat_2, recType_ref, by = c("recType.y" = "recType"))
  dat_2 <- subset(dat_2, !selectMeth.y %in% c("NotApp", "NotSam"))
  dat_21 <- summarise(group_by(dat_2, recType_name, stratumName.x, stratumName.y, selectMeth.y), no = length(selectMeth.y))
  
p <-  dat_21 %>%
  mutate(
    selectMeth.y = cell_spec(
      selectMeth.y,
      color = "white",
      background  = ifelse(selectMeth.y %in% c("NPAH", "NPEJ", "NPQS"), "red", "green"),
      bold = T
    )
  )  %>%
  kable(
    booktabs = T,
    col.names = c(
      "Sampling unit type",
      "Stratum above",
      "Stratum here",
      "Selection method",
      "Number of samples"
    ),
    escape = FALSE
  ) %>%
  kable_styling()

cat(p)
cat("  \n")
}

```


## None sampled
### Overall - summary


```{r, results = "asis"}

dat_3 <- data.frame(matrix(ncol = length(noSampReason$Key)+4 , nrow = 0))
colnames(dat_3) <- c(c("su","recType_name", "samp"), noSampReason$Key, c("No reason given"))

dat_3$su <- as.character(dat_3$su)
dat_3$recType_name <- as.character(dat_3$recType_name)
dat_3$samp <- as.character(dat_3$samp)
dat_3$`Industrial Decline` <- as.integer(dat_3$`Industrial Decline`)
dat_3$`No Answer` <- as.integer(dat_3$`No Answer`)
dat_3$`No Contact Details` <- as.integer(dat_3$`No Contact Details`)
dat_3$`Not Avalible` <- as.integer(dat_3$`Not Avalible`)
dat_3$`Observer Decline` <- as.integer(dat_3$`Observer Decline`)
dat_3$Other <- as.integer(dat_3$Other)
dat_3$`Quota Reached` <- as.integer(dat_3$`Quota Reached`)
dat_3$`No reason given` <- as.integer(dat_3$`No reason given`)


for (i in c(4:(length(h1_test)-1))) {
  
  dat_03 <- left_join(h1_test[[i]], recType_ref)
  dat_03$su <- gsub("su", "", dat_03$su)
  
  dat_03 <- mutate(dat_03, noSampReason = ifelse(is.na(noSampReason) & samp == "N", "No reason given", noSampReason))
  
  dat_031 <- summarise(group_by(dat_03, su, recType_name, samp, noSampReason), no = length(id))
  dat_031 <- filter(dat_031, !is.na(noSampReason))
  dat_031 <- spread(dat_031, key = noSampReason, value = no, fill = 0)
  dat_3 <- bind_rows(dat_3, dat_031)
}

heading <- distinct(left_join(h1_test$de, h1_test$sd, by = c("DEid" = "DEid")), ctry, DEyear, DEsampScheme)

summary_select <- bind_cols(heading, dat_3)

p <- summary_select %>%
  # mutate(
  #   Prob = cell_spec(
  #     Prob,
  #     color = ifelse(Prob > 0, "white", "white"),
  #     background  = ifelse(Prob > 0, "green", "red"),
  #     bold = ifelse(Prob > 0, T, T)
  #   ),
  #   None_prob = cell_spec(
  #     None_prob,
  #     color = ifelse(None_prob > 0, "white", "white"),
  #     background  = ifelse(None_prob > 0, "red", "green"),
  #     bold = ifelse(None_prob > 0, T, T)
  #   )
  # )  %>%
  kable(
    booktabs = T,
    col.names = c(c(
      "Sampling scheme",
      "Year",
      "Country",
      "Sampling unit level",
      "Sampling unit type",
      "Sampled or not"
    ), names(summary_select[,c(7:14)])),
    escape = FALSE
  ) %>%
  kable_styling()

if (nrow(subset(summary_select, samp =="N")) >1) {
  
  cat(p)
  
} else {
  
  cat("**Only sampled data declared**")
  
  cat(p)
}

```

### Detailed summary per sampling level

Number of none-samples per reason for not sampling. 

In each table the stratum from the level above is stated to make identification easier.

#### 1. sampling unit

```{r, results = "asis"}

dat_1 <- left_join(left_join(h1_test$de, h1_test$sd, by = c("DEid" = "DEid")), h1_test$su1, by = c("id" = "idAbove"))

dat_1 <- left_join(dat_1, recType_ref, by = c("recType.y" = "recType"))
dat_1 <- mutate(dat_1, noSampReason = ifelse(is.na(noSampReason) & samp == "No", "No reason given", noSampReason))
dat_11 <- summarise(group_by(dat_1, recType_name, DEsampScheme, DEyear, ctry, DEstratumName, stratumName, samp, noSampReason), no = length(selectMeth))


  dat_11 <- filter(dat_11, !is.na(noSampReason))


p <- dat_11 %>%
  # mutate(
  #   selectMeth = cell_spec(
  #     selectMeth,
  #     color = "white",
  #     background  = ifelse(selectMeth %in% c("NPAH", "NPEJ", "NPQS"), "red", "green"),
  #     bold = T
  #   )
  # )  %>%
  kable(
    booktabs = T,
    col.names = c(
      "Sampling unit type",
      "Sampling scheme",
      "Year",
      "Country",
      "Stratum above",
      "Stratum here",
      "Sampled or not",
      "Reason for not sampling",
      "Number of samples"
    ),
    escape = FALSE
  ) %>%
  kable_styling()

if (nrow(subset(dat_11, samp =="N")) >1) {
  
  cat(p)
  
} else {
  
  cat("**Only sampled data declared**")
  
  cat(p)
}


```

