---
title: "Simulated dataset for WKRDB_EST truc"
author: "Laurent Dubroca"
date: "`r format(Sys.time(),'%d/%m/%Y')`"
output: 
  pdf_document
---

# Introduction

- simulation of a fish population fished by a fleet of vessels. Three gears with
different selectivity curves. Then data are generated according to the sampling
scheme used in the french at-sea sampling scheme.
- based on the simulated data, a dirty (and probably buggy) hierarchy one set of
  files is generated.


```{r global_options, include=T,cache=F,echo=T,warning=F,message=F,progress=F,verbose=F}
#knitr option
operationnel<-TRUE
knitr::opts_chunk$set(echo=FALSE, 
		      warning=!operationnel, 
		      message=!operationnel,
		      fig.height=4,
		      progress=!operationnel,
		      verbose=!operationnel,
		      include=TRUE,dev='png',autodep=FALSE)
#package
library(dplyr)
library(ggplot2)
library(sampling)
library(FSA)
library(FSAsim)
library(purrr)
##library(openxlsx);library(mailR)
library(pander);library(captioner)
#local fct
#source("credo_fct.R")
#initialise les fonctions pour légender tables et figures
tabcap <- captioner(prefix = "Table")
figcap <- captioner(prefix = "Figure")

#general parameters
nbpopfish<-100000 #number of fish in the virtual population
natmortpopfish<-0.3 # the natural mortality of the virtual population
vbLinf <- 500  #von berta: Linf value of the virtual fish
vbK <- 0.2#von berta: K value of the virtual fish
vbt0 <- 0#von berta: t0 of the virtual fish
vbsigma <- 10#von berta sd around age for length distribution 
rtp<-data.frame(sciname="Virtus piscem",a=0.01/1000,b=3)

#fix randomness with the number of the Beast
set.seed(666)
```


# A virtual fish population 

Simulation of a virtual fish population.


```{r popsim1, include=T,cache=T,echo=T,eval=T}
#a population of fish A
popage<-simAges(N0=nbpopfish,A=natmortpopfish,incl.zero=T)
#their lengths
poplen <- simLenFromAge(popage,Linf=vbLinf,K=vbK,t0=vbK,sigma=vbsigma)
#the population dataframe (filtering out negative length if any) 
popfish <- data.frame(age=popage,len=poplen)%>%filter(len>0)
#adding weight
popfish$w<-rtp$a*(popfish$len/10)^rtp$b
#trunc the fish size
popfish$len <- lencat(popfish$len,w=10)
#some text summarizing the simulation parameters
txtparam1<-paste0("n=",nbpopfish,",A=",natmortpopfish)
txtparam2<-paste0("Linf=",vbLinf,",K=",vbK,",t0=",vbt0,",sd=",vbsigma)
txtparam3<-paste0("a=",rtp$a,",b=",rtp$b)#,",t0=",vbt0,",sd=",vbsigma)
```

```{r popsimplt1, include=T,cache=T,echo=T,eval=T,height=5}
p1<-ggplot(popfish,aes(x=age,y=len))+geom_point(alpha=.4)+
	theme_bw()+xlab("Age (year)")+ylab("Length (mm)")+
	ggtitle(paste0("Length at age of the simulated population\n",
		       txtparam1,"\n",txtparam2,"\n",txtparam3))
print(p1)
```

```{r popsimplt1cap,cache=T,eval=T,echo=T,results="asis"}
cat(figcap(name="popsimplt1",caption="Length at age in the virtual population"))
```

```{r popsimplt1bis, include=T,cache=T,echo=T,eval=T,height=5}
p1<-ggplot(popfish,aes(x=len,y=w))+geom_point(alpha=.4)+
	theme_bw()+xlab("Length (mm)")+ylab("Weight (kg)")+
	ggtitle(paste0("Weight at length of the simulated population\n",
		       txtparam1,"\n",txtparam2,"\n",txtparam3))
print(p1)
```

```{r popsimplt1biscap,cache=T,eval=T,echo=T,results="asis"}
cat(figcap(name="popsimplt1",caption="Weight at length in the virtual population"))
```

```{r popsimplt2, include=T,cache=T,echo=T,eval=T,height=5}
p2<-ggplot(tidyr::gather(popfish),aes(value))+geom_histogram()+#binwidth=c(1,10))+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distribution of the simulated population\n",txtparam1,"\n",txtparam2))
print(p2)
```
```{r popsimplt2cap,cache=T,eval=T,echo=FALSE,results="asis"}
cat(figcap(name="popsimplt2",caption="Length distribution of the virtual population"))
```


### A virtual fleet targeting a virtual fish population 

A gear is defined by its selectiviy curve. This curve is supplied using a beta
distribution.
A graphical representation of the selectivity curves is provided.

```{r fleetexample, include=T,cache=T,echo=T,eval=T}
selex1<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=10,shape2=20),gear="1")
selex2<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=4,shape2=3),gear="2")
selex3<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=8,shape2=1),gear="3")
selex<-rbind(selex1,selex2,selex3)
p1<-ggplot(selex,aes(x=x,y=sel,color=gear))+
	geom_path()+
	theme_bw()+ylab("Selectity")+xlab("Normalized size")+
	ggtitle(paste0("Selectivity curve from three gamma distributions"))#\n",txtparam1,"\n",txtparam2))
print(p1)
```

Population by gears + trips: 

```{r fleetandtrip, include=T,cache=T,echo=T,eval=T}
#population accessible to each trip
lengear1<-simLenSelectP(popfish$len, alpha = 10, beta = 20, max.height = 1, show = FALSE)
lengear2<-simLenSelectP(popfish$len, alpha = 4, beta = 3, max.height = 1, show = FALSE)
lengear3<-simLenSelectP(popfish$len, alpha = 8, beta = 1, max.height = 1, show = FALSE)
popfish1<-popfish%>%filter(lengear1)%>%mutate(gear="1")
popfish2<-popfish%>%filter(lengear2)%>%mutate(gear="2")
popfish3<-popfish%>%filter(lengear3)%>%mutate(gear="3")
popfishgear<-rbind(popfish1,popfish2,popfish3)
p1<-ggplot(tidyr::gather(popfishgear,key="key",value="value",age:len),
	   aes(value,fill=gear,group=gear))+
	geom_histogram(position="dodge")+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distributions available by gear"))
print(p1)

#generate some trips
#a fct to generate 1 haul, then 1 trip, then 1 vessel
do1haul<-function(df){
	sample_n(df,sample(50:450,1))%>%
		mutate(idhaul=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
do1trip<-function(df,nbhaul=sample(1:10,1)){
	do.call("rbind",rerun(nbhaul,do1haul(df)))%>%
		mutate(idtrip=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
do1vessel<-function(df,nbtrip=sample(1:50,1)){
	do.call("rbind",rerun(nbtrip,do1trip(df)))%>%
		mutate(idvessel=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
	

#generate 100 trip for popfish1
vess1<-do.call("rbind",rerun(50,do1vessel(popfish1))) %>%mutate(gear="1")
vess2<-do.call("rbind",rerun(30,do1vessel(popfish1))) %>%mutate(gear="2")
vess3<-do.call("rbind",rerun(10,do1vessel(popfish1))) %>%mutate(gear="3")
popvess<-rbind(vess1,vess2,vess3)%>%ungroup()%>%mutate(pk=row_number())
#select vessel and then trip and then systematic on haul 1 on 3
vesselection<-popvess%>%select(gear,idvessel)%>%distinct()%>%
	group_by(gear)%>%
	mutate(n=n_distinct(idvessel))%>%
	sample_frac(.1,weight=n)#unique(popvess$vessid)
#select n trip by vessel
tripselection<-popvess%>%select(idvessel,idtrip)%>%distinct()%>%
	filter(idvessel%in%vesselection$idvessel)%>%
	group_by(idvessel)%>%
	mutate(n=n_distinct(idtrip))%>%
	sample_n(1)
#select 1 haul every 3 haul
haulselection<-popvess%>%select(idvessel,idtrip,idhaul)%>%distinct()%>%
	filter(idtrip%in%tripselection$idtrip)%>%
	group_by(idtrip)%>%
	mutate(id=row_number())%>%
	filter(id%in%seq(1,10,2))%>%select(-id)
#select n length by haul and add a samp id
selvess<-left_join(haulselection%>%ungroup(),popvess)%>%
	group_by(idhaul)%>%
	sample_n(30)%>%mutate(idsamp=group_indices())
#pop file with sampling id
popvess<-full_join(popvess,selvess%>%select(pk,idsamp)%>%distinct())
saveRDS(popvess,file="./data/popvess.rds")
```


# Simulation to RDBES data

From the simulated data, a hierarchy 1 is generated (some tables are omitted).

```{r gendata, include=T,cache=T,echo=T,eval=T}

###################################################
#reverse methods
#start from lower hierarchy then build up information needed in each table
FM<-popvess%>%filter(!is.na(idsamp))%>%group_by(idsamp,FMclass=len)%>%
	summarise(FMnumberAtUnit=n())%>%
	ungroup()%>%
	transmute(FMid=row_number(),
		  SAid=idsamp,
		  FMrecType="FM",
		  FMclass,
		  FMnumberAtUnit,
		  FMtype="length",
		  FMMeasurementEquipement="",
		  FMaccuracy="",
		  Fmsampler="")

####################################################
#SA
tmp<-popvess%>%group_by(idhaul)%>%mutate(haulw=sum(w))%>%
	ungroup()%>%group_by(idsamp,haulw)%>%summarise(sampw=sum(w))%>%ungroup()
SA<-tmp%>%transmute(SAid=idsamp,
		  SAparentID=NA,
		  SSid="SSsppA",
		  SArecType="SA",
		  SAnatCode="",
		  SAstratification="N",
		  SAstratum="U",
		  SAsppCode="Genus species",
		  SAcommSpp="AAA",
		  SApres="",
		  SAcatchCat="Lan",
		  SAlandCat="HUC",
		  SAcommCatScl="",
		  SAcommCat="",
		  SAsex="U",
		  SAunitType="Haul",
		  SAtotalWtLive=haulw,
		  SAsampWtLive=sampw,
		  SAtotal=1,
		  SAsampled=1,
		  SAprob=NA,
		  SAselectMeth="Census",
		  SAlowHierarchy="FM",
		  SAsampler="Observer",
		  SAnoSampReasonFM="",
		  SAnoSampReasonBV="",
		  SAtotalWtMes=haulw,
		  SAsampWtMes=sampw,
		  SAconFacMesLive=1,
		  SAspecState="")

#species selection 
tmp<-popvess%>%filter(!is.na(idsamp))%>%select(idhaul)%>%distinct()
SS<-tmp%>%transmute(SSid="SSsppA",
	       LEid=NA,
	       FOid=idhaul,
	       SLid="SLsppA",
	       SSrecType="SS",
	       SSstratification="U",
	       SSobsActTyp="Haul",
	       SScatchCat="Lan",
	       SSobsTyp="visual",
	       SSstratum="U",
	       SSclustering="No",
	       SSclusterName="U",
	       SSsampler="",
	       SSsppListName="",
	       SStotal=1,
	       SSsampled=1,
	       SSselectMeth="Census",
	       SSselectMethCluster="U",
	       SStotalClusters="",
	       SSsampledClusters="",
	       SSprobCluster="",
	       SSnoSampReason="")
###########################################################
#FO
tmp<-popvess%>%group_by(idtrip)%>%mutate(check=sum(idsamp,na.rm=T))%>%
	ungroup()%>%filter(check>0)%>%
	select(idhaul,idtrip,idsamp)%>%distinct()%>%
	group_by(idtrip)%>%mutate(tothaul=n_distinct(idhaul),samphaul=n_distinct(idsamp)-1) %>%
	ungroup()%>%
	select(idhaul,idtrip,tothaul,samphaul)%>%distinct()%>%
	group_by(idtrip)%>%mutate(n=row_number())%>%ungroup()
FO<-tmp%>%transmute(FOid=idhaul,
	       FTid=idtrip,
	       SDid="",
	       FOrecType="FO",
	       FOstratification="N",
	       FOhaulNum=n,
	       FOstratum="U",
	       FOclustering="No",
	       FOclusterName="",
	       FOsampler="",
	       FOaggLev="H",
	       FOval="V",
	       FOcatReg="Lan",
	       FOstartDate="",
	       FOstartTime="",
	       FOendDate="",
	       FOendTime="",
	       FOdur="",
	       FOstartLat="",
	       FOstartLon="",
	       FOstopLat="",
	       FOstopLon="",
	       FOecoZone="",
	       FOarea="",
	       FOstatRect="",
	       FOsubRect="",
	       FOdep="",
	       FOwaterDep="",
	       FOnatCat="",
	       FOmetier5="",
	       FOmetier6="",
	       FOgear="",
	       FOmeshSize="",
	       FOselDev="",
	       FOselDevMeshSize="",
	       FOtarget="",
	       FOobsCo="Ha",
	       FOtotal=tothaul,
	       FOsampled=samphaul,
	       FOprob="",
	       FOselectMeth="SYSS",
	       FOselectMethCluster="",
	       FOtotalClusters="",
	       FOsampledClusters="",
	       FOprobCluster="",
	       FOnoSampReason="")

###########################################################
#FT
tmp<- popvess%>%group_by(idvessel)%>%mutate(nbtrip=n_distinct(idtrip))%>%
	group_by(idvessel,idtrip)%>%mutate(nbhaul=n_distinct(idhaul))%>%
	group_by(idvessel,idtrip,idhaul)%>%mutate(nbsamp=n_distinct(idsamp)-1)%>%
	ungroup()%>%
	select(idvessel,idtrip,nbtrip,nbhaul,nbsamp)%>%distinct() %>%
	filter(nbsamp>0)


FT<-tmp%>%transmute(FTid=idtrip,
	       OSid="",
	       VSid=idvessel,
	       VDid=idvessel,
	       SDid="",
	       FOid="",
	       TEid="",
	       FTrecType="FT",
	       FTnatCode="",
	       FTstratification="N",
	       FTstratum="U",
	       FTclustering="No",
	       FTclusterName="U",
	       FTsampler="",
	       FTfoNum=nbhaul,
	       FTdepLoc="",
	       FTdepDate="",
	       FTdepTime="",
	       FTarvLoc="",
	       FTarvDate="",
	       FTarvTime="",
	       FTtotal=nbtrip,
	       FTsampled=nbsamp,
	       FTprob="",
	       FTselectMeth="random",
	       FTselectMethCluster="",
	       FTtotalClusters="",
	       FTsampledClusters="",
	       FTprobCluster="",
	       FTnoSampReason=""
	       )

#VS
tmp<- popvess%>%group_by(gear)%>%mutate(nbvessel=n_distinct(idvessel))%>%
	group_by(gear,idvessel)%>%mutate(nbsamp=n_distinct(idsamp)-1)%>%
	ungroup()%>%
	filter(nbsamp>0)%>%
	select(gear,idvessel,nbvessel,nbsamp)%>%distinct()

VS<-tmp%>%transmute(VSid=idvessel,
		SDid=2,
		VDid=idvessel,
		TEid="",
		VSrecType="VS",
		VSstratification="Y",
		VSstratum=paste("Gear",gear),
		VSclustering="N",
		VSclusterName="U",
		VSsampler="Someone",
		VStotal=nbvessel,
		VSsampled=nbsamp,
		VSprob=.1,
		VSselectMeth="random",
		VSselectMethCluster="",
		VStotalClusters="",
		VSsampledClusters="",
		VSprobCluster="",
		VSnoSampReason="")
DE<-data.frame(DEid=1,
        DErecType="DE",
	DEsampScheme="sim H1",
	DEyear=2054,
	DEstratum=c("Stratified by fleet"),
	DEhierarchyCor=c("Yes"),
	DEhierarchy="H1")
SD<-data.frame(SDid=1,
	       DEid=1,
	       SDrecType="SD",
	       SDctry="XXX",
	       SDinst="Fishery Agency")


simH1<-list(DE,SD,VS,FT,FO,SS,SA,FM)
saveRDS(simH1,file="./data/simH1.rds")
```
