---
title: "Sampling strategies for age composition in commercial sampling and its effects on age distribution estimates: a simulation tool"
author: "Laurent Dubroca in WKBIOPTIM3"
date: "`r format(Sys.time(),'%d/%m/%Y')`"
output: 
  pdf_document
bibliography: '/home/moi/datahome/work/biblio/enfin/biblioloran.bib'
---

# Introduction

Some words about the framework, the needs and etc.
This document follows the framework proposed by @aanes2015.


```{r global_options, include=T,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
#knitr option
operationnel<-TRUE
knitr::opts_chunk$set(echo=FALSE, 
		      warning=!operationnel, 
		      message=!operationnel,
		      fig.height=4,
		      progress=!operationnel,
		      verbose=!operationnel,
		      include=TRUE,dev='png',autodep=FALSE)
#package
library(dplyr)
library(ggplot2)
library(sampling)
library(FSA)
library(FSAsim)
library(purrr)
##library(openxlsx);library(mailR)
library(pander);library(captioner)
#local fct
#source("credo_fct.R")
#initialise les fonctions pour légender tables et figures
tabcap <- captioner(prefix = "Table")
figcap <- captioner(prefix = "Figure")

#general parameters
nbpopfish<-100000 #number of fish in the virtual population
natmortpopfish<-0.3 # the natural mortality of the virtual population
vbLinf <- 500  #von berta: Linf value of the virtual fish
vbK <- 0.2#von berta: K value of the virtual fish
vbt0 <- 0#von berta: t0 of the virtual fish
vbsigma <- 10#von berta sd around age for length distribution 
rtp<-data.frame(sciname="Virtus piscem",a=0.01/1000,b=3)

#fix randomness with the number of the Beast
set.seed(666)
```

# Materials and methods


## Simulation

Simulation of a virtual fish population.
Principles and paramaters.


### A virtual fish population 

Comments of `r figcap("popsimplt1",display="cite")` and `r figcap("popsimplt2",display="cite")`.


```{r popsim1, include=T,cache=T,echo=F,eval=T}
#a population of fish A
popage<-simAges(N0=nbpopfish,A=natmortpopfish,incl.zero=T)
#their lengths
poplen <- simLenFromAge(popage,Linf=vbLinf,K=vbK,t0=vbK,sigma=vbsigma)
#the population dataframe (filtering out negative length if any) 
popfish <- data.frame(age=popage,len=poplen)%>%filter(len>0)
#adding weight
popfish$w<-rtp$a*(popfish$len/10)^rtp$b
#trunc the fish size
popfish$len <- lencat(popfish$len,w=10)
#some text summarizing the simulation parameters
txtparam1<-paste0("n=",nbpopfish,",A=",natmortpopfish)
txtparam2<-paste0("Linf=",vbLinf,",K=",vbK,",t0=",vbt0,",sd=",vbsigma)
txtparam3<-paste0("a=",rtp$a,",b=",rtp$b)#,",t0=",vbt0,",sd=",vbsigma)
```

```{r popsimplt1, include=T,cache=T,echo=F,eval=T,height=5}
p1<-ggplot(popfish,aes(x=age,y=len))+geom_point(alpha=.4)+
	theme_bw()+xlab("Age (year)")+ylab("Length (mm)")+
	ggtitle(paste0("Length at age of the simulated population\n",
		       txtparam1,"\n",txtparam2,"\n",txtparam3))
print(p1)
ggsave("fishpop.png")
```

```{r popsimplt1cap,cache=T,eval=T,echo=FALSE,results="asis"}
cat(figcap(name="popsimplt1",caption="Length at age in the virtual population"))
```

```{r popsimplt1bis, include=T,cache=T,echo=F,eval=T,height=5}
p1<-ggplot(popfish,aes(x=len,y=w))+geom_point(alpha=.4)+
	theme_bw()+xlab("Length (mm)")+ylab("Weight (kg)")+
	ggtitle(paste0("Weight at length of the simulated population\n",
		       txtparam1,"\n",txtparam2,"\n",txtparam3))
print(p1)
```

```{r popsimplt1biscap,cache=T,eval=T,echo=FALSE,results="asis"}
cat(figcap(name="popsimplt1",caption="Weight at length in the virtual population"))
```

```{r popsimplt2, include=T,cache=T,echo=F,eval=T,height=5}
p2<-ggplot(tidyr::gather(popfish),aes(value))+geom_histogram()+#binwidth=c(1,10))+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distribution of the simulated population\n",txtparam1,"\n",txtparam2))
print(p2)
ggsave("fishpop2.png")
```
```{r popsimplt2cap,cache=T,eval=T,echo=FALSE,results="asis"}
cat(figcap(name="popsimplt2",caption="Length distribution of the virtual population"))
```


### A virtual fleet targeting a virtual fish population 

Principles.
A gear is define by its selectiviy curve. This curve is supplied using a beta
distribution.
An example of a selectivity curve

```{r fleetexample, include=T,cache=T,echo=F,eval=T}
selex1<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=10,shape2=20),gear="1")
selex2<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=4,shape2=3),gear="2")
selex3<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=8,shape2=1),gear="3")
selex<-rbind(selex1,selex2,selex3)
p1<-ggplot(selex,aes(x=x,y=sel,color=gear))+
	geom_path()+
	theme_bw()+ylab("Selectity")+xlab("Normalized size")+
	ggtitle(paste0("Selectivity curve from three gamma distributions"))#\n",txtparam1,"\n",txtparam2))
print(p1)
ggsave("vesselsel.png")
```

Population by gears + trips: 

```{r fleetandtrip, include=T,cache=T,echo=F,eval=T}
#population accessible to each trip
lengear1<-simLenSelectP(popfish$len, alpha = 10, beta = 20, max.height = 1, show = FALSE)
lengear2<-simLenSelectP(popfish$len, alpha = 4, beta = 3, max.height = 1, show = FALSE)
lengear3<-simLenSelectP(popfish$len, alpha = 8, beta = 1, max.height = 1, show = FALSE)
popfish1<-popfish%>%filter(lengear1)%>%mutate(gear="1")
popfish2<-popfish%>%filter(lengear2)%>%mutate(gear="2")
popfish3<-popfish%>%filter(lengear3)%>%mutate(gear="3")
popfishgear<-rbind(popfish1,popfish2,popfish3)
p1<-ggplot(tidyr::gather(popfishgear,key="key",value="value",age:len),
	   aes(value,fill=gear,group=gear))+
	geom_histogram(position="dodge")+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distributions available by gear"))
print(p1)
ggsave("fish4vessel.png")

#generate some trips
#a fct to generate 1 haul, then 1 trip, then 1 vessel
do1haul<-function(df){
	sample_n(df,sample(50:450,1))%>%
		mutate(idhaul=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
do1trip<-function(df,nbhaul=sample(1:10,1)){
	do.call("rbind",rerun(nbhaul,do1haul(df)))%>%
		mutate(idtrip=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
do1vessel<-function(df,nbtrip=sample(1:50,1)){
	do.call("rbind",rerun(nbtrip,do1trip(df)))%>%
		mutate(idvessel=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
	

#generate 100 trip for popfish1
vess1<-do.call("rbind",rerun(50,do1vessel(popfish1))) %>%mutate(gear="1")
vess2<-do.call("rbind",rerun(30,do1vessel(popfish1))) %>%mutate(gear="2")
vess3<-do.call("rbind",rerun(10,do1vessel(popfish1))) %>%mutate(gear="3")
popvess<-rbind(vess1,vess2,vess3)%>%ungroup()%>%mutate(pk=row_number())
saveRDS(popvess,file="popvess.rds")
#select vessel and then trip and then systematic on haul 1 on 3
vesselection<-popvess%>%select(gear,idvessel)%>%distinct()%>%
	group_by(gear)%>%
	mutate(n=n_distinct(idvessel))%>%
	sample_frac(.1,weight=n)#unique(popvess$vessid)
#select n trip by vessel
tripselection<-popvess%>%select(idvessel,idtrip)%>%distinct()%>%
	filter(idvessel%in%vesselection$idvessel)%>%
	group_by(idvessel)%>%
	mutate(n=n_distinct(idtrip))%>%
	sample_n(1)
#select 1 haul every 3 haul
haulselection<-popvess%>%select(idvessel,idtrip,idhaul)%>%distinct()%>%
	filter(idtrip%in%tripselection$idtrip)%>%
	group_by(idtrip)%>%
	mutate(id=row_number())%>%
	filter(id%in%seq(1,10,2))%>%select(-id)
#select n length by haul and add a samp id
selvess<-left_join(haulselection%>%ungroup(),popvess)%>%
	group_by(idhaul)%>%
	sample_n(30)%>%mutate(idsamp=group_indices())
#pop file with sampling id
popvess<-full_join(popvess,selvess%>%select(pk,idsamp)%>%distinct())
saveRDS(popvess,file="popvess.rds")



if(F){
#using sampling to get a complex statified example
pipo<-popvess[order(popvess$gear,popvess$idvessel,popvess$idtrip,popvess$idhaul),c("gear","idvessel","idtrip","idhaul","w","len","age")]%>%transmute(gear=as.numeric(gear),idvessel=as.numeric(idvessel),w)%>%distinct()
mstage(pipo,stage=list("stratified",""),varnames=list("gear","idvessel"),size=list(c(3,3),method=list("srswor","srswor"))
pipo<-data.frame(gear=c(rep(1,5),rep(2,3)),idvessel=1:8,w=rnorm(8,mean=10))
strata(pipo,c("gear"),size=c(2,3),method="srswor")
mstage(pipo,stage=list("stratified","stratified"),
       varnames=list("gear","idvessel"),
       size=list(2,1),method=c("srswor"))
mstage(pipo,stage=list("stratified",""),
       varnames=list("gear","idvessel"),
       size=list(c(2,1),c(1,1)),method=c("","srswor"))
}


###################################################
#reverse methods
#start from lower hierarchy then build up information needed in each table
FM<-popvess%>%filter(!is.na(idsamp))%>%group_by(idsamp,FMclass=len)%>%
	summarise(FMnumberAtUnit=n())%>%
	ungroup()%>%
	transmute(FMid=row_number(),
		  SAid=idsamp,
		  FMrecType="FM",
		  FMclass,
		  FMnumberAtUnit,
		  FMtype="length",
		  FMMeasurementEquipement="",
		  FMaccuracy="",
		  Fmsampler="")

####################################################
#SA
tmp<-popvess%>%group_by(idhaul)%>%mutate(haulw=sum(w))%>%
	ungroup()%>%group_by(idsamp,haulw)%>%summarise(sampw=sum(w))%>%ungroup()
SA<-tmp%>%transmute(SAid=idsamp,
		  SAparentID=NA,
		  SSid="SSsppA",
		  SArecType="SA",
		  SAnatCode="",
		  SAstratification="N",
		  SAstratum="U",
		  SAsppCode="Genus species",
		  SAcommSpp="AAA",
		  SApres="",
		  SAcatchCat="Lan",
		  SAlandCat="HUC",
		  SAcommCatScl="",
		  SAcommCat="",
		  SAsex="U",
		  SAunitType="Haul",
		  SAtotalWtLive=haulw,
		  SAsampWtLive=sampw,
		  SAtotal=1,
		  SAsampled=1,
		  SAprob=NA,
		  SAselectMeth="Census",
		  SAlowHierarchy="FM",
		  SAsampler="Observer",
		  SAnoSampReasonFM="",
		  SAnoSampReasonBV="",
		  SAtotalWtMes=haulw,
		  SAsampWtMes=sampw,
		  SAconFacMesLive=1,
		  SAspecState="")

#species selection 
tmp<-popvess%>%filter(!is.na(idsamp))%>%select(idhaul)%>%distinct()
SS<-tmp%>%transmute(SSid="SSsppA",
	       LEid=NA,
	       FOid=idhaul,
	       SLid="SLsppA",
	       SSrecType="SS",
	       SSstratification="U",
	       SSobsActTyp="Haul",
	       SScatchCat="Lan",
	       SSobsTyp="visual",
	       SSstratum="U",
	       SSclustering="No",
	       SSclusterName="U",
	       SSsampler="",
	       SSsppListName="",
	       SStotal=1,
	       SSsampled=1,
	       SSselectMeth="Census",
	       SSselectMethCluster="U",
	       SStotalClusters="",
	       SSsampledClusters="",
	       SSprobCluster="",
	       SSnoSampReason="")
###########################################################
#FO
tmp<-popvess%>%group_by(idtrip)%>%mutate(check=sum(idsamp,na.rm=T))%>%
	ungroup()%>%filter(check>0)%>%
	select(idhaul,idtrip,idsamp)%>%distinct()%>%
	group_by(idtrip)%>%mutate(tothaul=n_distinct(idhaul),samphaul=n_distinct(idsamp)-1) %>%
	ungroup()%>%
	select(idhaul,idtrip,tothaul,samphaul)%>%distinct()%>%
	group_by(idtrip)%>%mutate(n=row_number())%>%ungroup()
FO<-tmp%>%transmute(FOid=idhaul,
	       FTid=idtrip,
	       SDid="",
	       FOrecType="FO",
	       FOstratification="N",
	       FOhaulNum=n,
	       FOstratum="U",
	       FOclustering="No",
	       FOclusterName="",
	       FOsampler="",
	       FOaggLev="H",
	       FOval="V",
	       FOcatReg="Lan",
	       FOstartDate="",
	       FOstartTime="",
	       FOendDate="",
	       FOendTime="",
	       FOdur="",
	       FOstartLat="",
	       FOstartLon="",
	       FOstopLat="",
	       FOstopLon="",
	       FOecoZone="",
	       FOarea="",
	       FOstatRect="",
	       FOsubRect="",
	       FOdep="",
	       FOwaterDep="",
	       FOnatCat="",
	       FOmetier5="",
	       FOmetier6="",
	       FOgear="",
	       FOmeshSize="",
	       FOselDev="",
	       FOselDevMeshSize="",
	       FOtarget="",
	       FOobsCo="Ha",
	       FOtotal=tothaul,
	       FOsampled=samphaul,
	       FOprob="",
	       FOselectMeth="SYSS",
	       FOselectMethCluster="",
	       FOtotalClusters="",
	       FOsampledClusters="",
	       FOprobCluster="",
	       FOnoSampReason="")

###########################################################
#FT
tmp<- popvess%>%group_by(idvessel)%>%mutate(nbtrip=n_distinct(idtrip))%>%
	group_by(idvessel,idtrip)%>%mutate(nbhaul=n_distinct(idhaul))%>%
	group_by(idvessel,idtrip,idhaul)%>%mutate(nbsamp=n_distinct(idsamp)-1)%>%
	ungroup()%>%
	select(idvessel,idtrip,nbtrip,nbhaul,nbsamp)%>%distinct() %>%
	filter(nbsamp>0)


FT<-tmp%>%transmute(FTid=idtrip,
	       OSid="",
	       VSid=idvessel,
	       VDid=idvessel,
	       SDid="",
	       FOid="",
	       TEid="",
	       FTrecType="FT",
	       FTnatCode="",
	       FTstratification="N",
	       FTstratum="U",
	       FTclustering="No",
	       FTclusterName="U",
	       FTsampler="",
	       FTfoNum=nbhaul,
	       FTdepLoc="",
	       FTdepDate="",
	       FTdepTime="",
	       FTarvLoc="",
	       FTarvDate="",
	       FTarvTime="",
	       FTtotal=nbtrip,
	       FTsampled=nbsamp,
	       FTprob="",
	       FTselectMeth="random",
	       FTselectMethCluster="",
	       FTtotalClusters="",
	       FTsampledClusters="",
	       FTprobCluster="",
	       FTnoSampReason=""
	       )

#VS
tmp<- popvess%>%group_by(gear)%>%mutate(nbvessel=n_distinct(idvessel))%>%
	group_by(gear,idvessel)%>%mutate(nbsamp=n_distinct(idsamp)-1)%>%
	ungroup()%>%
	filter(nbsamp>0)%>%
	select(gear,idvessel,nbvessel,nbsamp)%>%distinct()

VS<-tmp%>%transmute(VSid=idvessel,
		SDid=2,
		VDid=idvessel,
		TEid="",
		VSrecType="VS",
		VSstratification="Y",
		VSstratum=paste("Gear",gear),
		VSclustering="N",
		VSclusterName="U",
		VSsampler="Someone",
		VStotal=nbvessel,
		VSsampled=nbsamp,
		VSprob=.1,
		VSselectMeth="random",
		VSselectMethCluster="",
		VStotalClusters="",
		VSsampledClusters="",
		VSprobCluster="",
		VSnoSampReason="")


all<-list(VS,FT,FO,SS,SA,FM)

#inclusion proba basique
pipo<-data.frame(sample(1:4,4))

inclusionprobabilities(c(1,1,1,1),2)
inclusionprobabilities(c(2,2,5,2),2)


stop()
#fct rdbes to pik
RDBESmodel<-readRDS("RDBESmodel.rds")
rname<-RDBESmodel%>%select(shortname,R.Name)

rname%>%filter(grepl("cluster",tolower(R.Name)))
rname%>%filter(grepl("strat",tolower(R.Name)))
rname%>%filter(grepl("selectmeth",tolower(R.Name)))
rname%>%filter(grepl("prob",R.Name))

#extract info



#build table
#H1: DE/SD/VD/VS/FT/FO/SS/SL/SA/FM/BV
tabstruc<-readRDS("RDBESmodel.rds")
idname<-which(grepl("ID",tabstruc$Short.Description))
#DE: 4 design here
#4 design
DE<-data.frame(DEid=1:2,
        DErecType="DE",
	DEsampScheme=paste("Program",1:2),
	DEyear=2054,
	DEstratum=c("Total census","Stratified by fleet")
	DEhierarchyCor=c("Yes"),
	DEhierarchy="H1")
#SD: sampling details 
SD<-data.frame(SDid=1:2,
	       DEid=1:2,
	       SDrecType="SD",
	       SDctry="EU",
	       SDinst="Fishery Agency")
#VD: vessel details
VD<-data.frame(VDid=1:length(unique(popvess$idvess)),
	       VDrecType="VD",
	       VDencrCode=sort(unique(popvess$idvess)),
	       VDhomePort="somewhere",
	       VDflgCtry="EU",
	       VDlen=25,
	       VDlenCat="25",
	       VDpwr=1000,
	       VDsize=100,
	       VDsizeUnit="GT",
	       VDtype="other boats")
#VS: vessel selection : 2 type census vs stratified sampling
#1 sensus
tmp<-popvess%>%select(idvessel,idtrip)%>%distinct()%>%
	group_by(idvessel)%>%summarise(n=n())%>%
	ungroup()%>%mutate(pi=n/sum(n))%>%ungroup()
VS1<-data.frame(VSid=1:nrow(tmp),
		SDid=1,
		VDid=tmp$idvessel,
		TEid=NA,
		VSrecType="VS",
		VSstratification="N",
		VSstratum="Complete fleet",
		VSclustering="N",
		VSclusterName="U",
		VSsampler="Someone",
		VStotal=length(unique(popvess$idvessel)),
		VSsampled=length(unique(popvess$idvessel)),
		VSprob=NA,
		VSselectMeth="Census",
		VSselectMethCluster="",
		VStotalClusters=NA,
		VSsampledClusters=NA,
		VSprobCluster=NA,
		VSnoSampReason="")
#2 stratification
tmp<-selvess%>%select(idvessel,idtrip)%>%distinct()%>%
	group_by(idvessel)%>%summarise(n=n())%>%
	ungroup()%>%mutate(pi=n/sum(n))%>%ungroup()
tmp2<-popvess%>%select(idvessel,gear)%>%distinct()%>%
	group_by(gear)%>%mutate(ntot=n_distinct(idvessel))%>%ungroup()
tmp<-left_join(tmp,tmp2)%>%arrange(idvessel)%>%group_by(gear)%>%
	mutate(nsamp=n_distinct(idvessel))%>%ungroup()%>%
	mutate(pi2=nsamp/ntot)
VS2<-data.frame(VSid=1:nrow(tmp),
		SDid=2,
		VDid=tmp$idvessel,
		TEid=NA,
		VSrecType="VS",
		VSstratification="Y",
		VSstratum="Stratified by gear",
		VSclustering="Y",
		VSclusterName=paste("Gear",tmp$gear),
		VSsampler="Someone",
		VStotal=tmp$ntot,
		VSsampled=tmp$nsamp,
		VSprob=NA,
		VSselectMeth="Random",
		VSselectMethCluster="Systematic",
		VStotalClusters=3,
		VSsampledClusters=3,
		VSprobCluster=1/3,
		VSnoSampReason="")
#VS total
VS<-rbind(VS1,VS2)

#H1: DE/SD/VD/VS/FT/FO/SS/SL/SA/FM/BV
#FT: Fishing trip
tmp<-selvess%>%select(idvessel,idtrip)%>%distinct()%>%
	group_by(idvessel)%>%summarise(n=n())%>%
	ungroup()%>%mutate(pi=n/sum(n))%>%ungroup()
tmp2<-popvess%>%select(idvessel,gear)%>%distinct()%>%
	group_by(gear)%>%mutate(ntot=n_distinct(idvessel))%>%ungroup()
tmp<-left_join(tmp,tmp2)%>%arrange(idvessel)%>%group_by(gear)%>%
	mutate(nsamp=n_distinct(idvessel))%>%ungroup()%>%
	mutate(pi2=nsamp/ntot)

FT1<-data.frame(FTid=,
		OSid=NA,
		VSid=,
		VDid=,
		SDid
		FOid
		TEid
		FTrecType="FT",
		FTnatCode="EU",
		FTstratification
		FTstratum
		FTclustering
		FTclusterName
		FTsampler
		FTfoNum
		FTdepLoc
		FTdepDate
		FTdepTime
		FTarvLoc
		FTarvDate
		FTarvTime
		FTtotal
		FTsampled
		FTprob
		FTselectMeth
		FTselectMethCluster
		FTtotalClusters
		FTsampledClusters
		FTprobCluster
		FTnoSampReason
		)



	       	
tmp<-tabstruc%>%filter(tabname=="Vessel Selection")
dattmp<-data.frame(matrix(nrow=0,ncol=nrow(tmp)))
names(dattmp)<-tmp$R.Name
str(dattmp)

#length distribution of the trips by gear
p1<-ggplot(tidyr::gather(tripgear,key="key",value="value",age:len),
	   aes(value,fill=gear,group=gear))+
	#geom_histogram(alpha=1,position="dodge",binwidth=10)+
	geom_histogram(alpha=1,position="dodge")+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distributions of the trips by gear"))
print(p1)
#raise the trip data to pop using weight
wpopgear<-popfishgear%>%group_by(gear)%>%summarise(wtot=sum(w))%>%ungroup()
wtripgear<-tripgear%>%group_by(gear,id)%>%mutate(wtrip=sum(w))%>%ungroup()%>%
		left_join(wpopgear,by="gear")

wgear2%>%group_by(gear)%>%summarise(w1=sum(wtrip))

wtripgear%>%select(gear,id,wtrip,wtot)%>%
	distinct()%>%group_by(gear)%>%
	summarise(wtottrip=sum(wtrip),wtot=unique(wtot))%>%
	group_by(gear)%>%
	mutate(f=wtot/wtottrip)%>%
	mutate(westim=f*wtottrip)%>%left_join(wpopgear)

tripgear%>%group_by(gear,id)%>%mutate(wtrip=sum(w))%>%
	group_by(gear)%>%mutate(wtottrip=sum(w))%>%
	left_join(wpopgear,by="gear")%>%
	select(wtottrip,wtot)%>%distinct()
	





#total age distribution for gear vs trip gear
pipo1<-popfishgear%>%group_by(age)%>%summarise(n=n())%>%ungroup()%>%
	mutate(type="popfishgear")
prop<-data.frame(gear=as.character(1:3),
		 prop=tapply(popfishgear$age,popfishgear$gear,sum)/300)
pipo2<-tripgear%>%
	left_join(prop,by="gear")%>%
	group_by(age,prop)%>%summarise(n=n())%>%ungroup()%>%
	mutate(type="popfishgear")


p1<-ggplot(gather(tripgear,key="key",value="value",age:len),
	   aes(value,fill=gear,group=gear))+
	#geom_histogram(alpha=1,position="dodge",binwidth=10)+
	geom_histogram(alpha=1,position="dodge")+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distributions of the trips by gear"))
print(p1)
```

## ALK sampling

Samples of ALK.
- samples the trips,
- independent samples,
- ??

```{r alk, include=T,cache=T,echo=F,eval=F}
# random sample of age from the total pop
ranalk<-function(nb=100,popfish,popfishgear){
	sampalk<-sample_n(popfish,100)
	#complete the alkprop with zero for the non sample length
	lenall<-data.frame(len=sort(unique(popfishgear$len)))
	sampalkall<-left_join(lenall,sampalk,by="len")#%>%
	sampalkall[is.na(sampalkall)]<-0
	#compute the alk
	alk <- xtabs(~len+age,data=sampalkall)#%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	#raise length to the pop using number
	rez1<-alkIndivAge(alkprop,age~len,data=popfishgear)#%>%mutate(trueage=age))
	rez1%>%mutate(id="ranalk",nb,rep=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
#generate 100 trip for popfish1
ranalk50<-do.call("rbind",rerun(30,ranalk(50,popfish,tripgear)))
ranalk100<-do.call("rbind",rerun(30,ranalk(100,popfish,tripgear)))
ranalk300<-do.call("rbind",rerun(30,ranalk(300,popfish,tripgear)))
ranalkall<-rbind(ranalk50,ranalk100,ranalk300)

# stratified random sample of age from the total pop
ranstratalk<-function(nb=5,popfish,popfishgear){
	nbbyage<-rep(5,length(unique(popfish$len)))
	sampalk<-strata(popfish,stratanames="len",size=nbbyage,method="srswr")
	sampalk<-popfish[sampalk$ID_unit,]
	#complete the alkprop with zero for the non sample length
	lenall<-data.frame(len=sort(unique(popfishgear$len)))
	sampalkall<-left_join(lenall,sampalk,by="len")#%>%
	sampalkall[is.na(sampalkall)]<-0
	#compute the alk
	alk <- xtabs(~len+age,data=sampalkall)#%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	#raise length to the pop using number
	rez1<-alkIndivAge(alkprop,age~len,data=popfishgear)#%>%mutate(trueage=age))
	rez1%>%mutate(id="ranstratalk",nb,rep=sub("\\.","",format(Sys.time(),"%s%OS3")))
}
#generate 100 trip for popfish1
ranstratalk1<-do.call("rbind",rerun(30,ranstratalk(1,popfish,tripgear)))
ranstratalk3<-do.call("rbind",rerun(30,ranstratalk(3,popfish,tripgear)))
ranstratalk5<-do.call("rbind",rerun(30,ranstratalk(5,popfish,tripgear)))
ranstratalkall<-rbind(ranstratalk1,ranstratalk3,ranstratalk5)

#summarize results
rezestim<-rbind(ranalkall,ranstratalkall)%>%group_by(id,age,gear,nb,rep)%>%
	summarise(nage=n())%>%
	group_by(id,gear,nb,age)%>%
	summarise(mage=mean(nage),sdage=sd(nage))%>%
	#mutate(nb=as.character(nb))%>%
	ungroup()

#add tripgear age real
pipo<-tripgear%>%group_by(gear,age)%>%
	summarise(mage=n())%>%ungroup()%>%
	transmute(id="ranalk",gear,nb="pop",age,mage,sdage=0)
rezestim<-rbind(rezestim,pipo)

#raising to the pop ?



p1<- ggplot(rezestim, aes(x=age, y=mage,color=nb))+
	  geom_path(alpha=.5)+
  facet_grid(id~gear,scale="free")#+theme_bw() 
  print(p1)


pipo<-rbind(ranalkall%>%select(age)%>%mutate(type="estim"),
	    tripgear%>%select(age)%>%mutate(type="pop"))
p1<-ggplot(pipo,aes(age,fill=type,group=type))+
	geom_histogram(alpha=1,position="dodge",binwidth=1)+
	theme_bw()+ylab("")+xlab("Length (mm)")+
	ggtitle(paste0("Length distribution of the trips by gear"))
print(p1)

earthmovdist::emdL1(rez1$age,popfishgear$age)


```



```{r data_and_param000, include=T,cache=T,echo=F,eval=F}





#apply age length key to length data
#from FSA
rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
rez1hist<-rez1%>%group_by(age)%>%summarise(n=n()*2,type="sampalk")
rez2hist<-rez1%>%group_by(age=trueage)%>%summarise(n=n()*2,type="sampreal")
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")
rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
ggplot(rez1hist,aes(x=age,y=n,color=type))+geom_point()+geom_path()

#a more complete example
#a population of fish
tabage<-simAges(N0=100000,A=.35,incl.zero=T)
#their lengths
tablen <- simLenFromAge(tabage,Linf=228,K=0.206,t0=0,sigma=8)
## Some summaries
df <- data.frame(age=tabage,len=tablen)%>%filter(len>0)
Summarize(len~age,data=df,digits=1)
plot(len~age,data=df,pch=16,col=rgb(0,0,0,0.1),xlab="Age",ylab="Length")
hist(df$len,n=50)





# independant sample
nbbyage<-rep(5,length(unique(df1$lenCat)))
df2<-strata(df1,stratanames="lenCat",size=nbbyage,method="srswr")
df2$trueage<-df1$trueage[df2$ID_unit]
alk <- xtabs(~lenCat+age,data=df2%>%transmute(age=trueage,lenCat))
alkprop <- prop.table(alk,margin=1)
round(alkprop,2)

```



```{r data_and_param, include=T,cache=T,echo=F,eval=F}





#apply age length key to length data
#from FSA
rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
rez1hist<-rez1%>%group_by(age)%>%summarise(n=n()*2,type="sampalk")
rez2hist<-rez1%>%group_by(age=trueage)%>%summarise(n=n()*2,type="sampreal")
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")
rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
ggplot(rez1hist,aes(x=age,y=n,color=type))+geom_point()+geom_path()

#a more complete example
#a population of fish
tabage<-simAges(N0=100000,A=.35,incl.zero=T)
#their lengths
tablen <- simLenFromAge(tabage,Linf=228,K=0.206,t0=0,sigma=8)
## Some summaries
df <- data.frame(age=tabage,len=tablen)%>%filter(len>0)
Summarize(len~age,data=df,digits=1)
plot(len~age,data=df,pch=16,col=rgb(0,0,0,0.1),xlab="Age",ylab="Length")
hist(df$len,n=50)
#1 gear no selectivity
## take some length sample randomly
df1<-df%>%sample_frac(0.01)
df1$lenCat <- lencat(df1$len,w=10)
df1<-df1%>%mutate(trueage=age)%>%select(-age)
## sample age based on df1
fct1<-function(df1,nb=5){
	nbbyage<-rep(nb,length(unique(df1$lenCat)))
	df2<-strata(df1,stratanames="lenCat",size=nbbyage,method="srswr")
	df2$trueage<-df1$trueage[df2$ID_unit]
	alk <- xtabs(~lenCat+age,data=df2%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
	rez1%>%mutate(n=nb)
}
#generate differente sample rate
rezall<-data.frame()
for(i in 1:30){
	rez0<-fct1(df1,i)
	rezall<-rbind(rez0,rezall)
}
#apply age length key to length data
#from FSA
prop<-1/0.01
rez1hist<-rezall%>%group_by(age,nbsamp=n)%>%summarise(n=n()*prop,type="sampalk")%>%ungroup()
rez2hist<-df1%>%group_by(age=trueage)%>%summarise(n=n()*prop,type="sampreal")%>%ungroup()
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")%>%ungroup()
#rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
#ggplot(rez1hist,aes(x=age,y=n,color=nbsamp))+geom_boxplot()+
ggplot(data=rez1hist,aes(x=age,y=n,group=age))+geom_boxplot()+
geom_point(data=rez1hist,aes(x=age,y=n,color=nbsamp,group=NULL),alpha=.2)+#geom_point(alpha=.2)+
	geom_line(data=rez2hist,aes(x=age,y=n,group=NULL),col="red")+
	geom_path(data=rez3hist,aes(x=age,y=n,group=NULL),col="green")

#######################################################
#a more complete example, again, with PSU and so on
#a population of fish
tabage<-simAges(N0=100000,A=.35,incl.zero=T)
#their lengths
tablen <- simLenFromAge(tabage,Linf=228,K=0.206,t0=0,sigma=8)
## Some summaries
df <- data.frame(age=tabage,len=tablen)%>%filter(len>0)
Summarize(len~age,data=df,digits=1)
plot(len~age,data=df,pch=16,col=rgb(0,0,0,0.1),xlab="Age",ylab="Length")
hist(df$len,n=50)
dfhist<-df%>%group_by(len)%>%summarise(n=n(),type="pop")

#1 gear with a given selectivity
aa<-simLenSelectP(df$len, alpha = 3, beta = 10, max.height = 1, show = FALSE)
df1<-df%>%filter(aa)%>%sample_frac(0.01)
dfhist1<-df1%>%group_by(len)%>%summarise(n=n(),type="gear1")
ggplot(rbind(dfhist,dfhist1),aes(x=len,y=n,col="type"))+geom_path()+facet_wrap(~type,scale="free")

#generate xx trips
dontrip<-function(df,n=5,prop0=0.01,alpha0=5,beta0=10){
	dotrip<-function(df,prop=prop0,alpha=alpha0,beta=beta0){
		aa<-simLenSelectP(df$len, alpha, beta, max.height = 1, show = FALSE)
		df%>%filter(aa)%>%sample_frac(0.01)%>%
			mutate(trip=paste(sample(letters,5),collapse=""))
	}
	aa<-rerun(n,dotrip(df,prop=prop0,alpha=alpha0,beta=beta0))
	aa<-do.call("rbind",aa)
	aa$lenCat <- lencat(aa$len,w=10)
	return(aa)
}

trip1<-dontrip(df,n=10,alpha0=2,beta0=7)
trip1hist<-trip1%>%group_by(len,trip)%>%summarise(n=n(),type="gear1")
prop<-max(trip1hist$n)/max(dfhist$n)
ggplot(trip1hist,aes(x=len,y=n,col=trip))+geom_path()+
	geom_path(data=dfhist,aes(x=len,y=n*prop,col="type"),col="black")

trip2<-dontrip(df,n=10,alpha0=6,beta0=7)
trip1hist<-trip2%>%group_by(len,trip)%>%summarise(n=n(),type="gear1")
prop<-max(trip1hist$n)/max(dfhist$n)
ggplot(trip1hist,aes(x=len,y=n,col=trip))+geom_path()+
	geom_path(data=dfhist,aes(x=len,y=n*prop,col="type"),col="black")

## sample age based on df1
fct1<-function(df1=rbind(trip1,trip2),nb=5){
df1=rbind(trip1,trip2);nb<-5
	nbbyage<-rep(nb,length(unique(df1$lenCat)))
	df2<-strata(df1,stratanames="lenCat",size=nbbyage,method="srswr")
	df2$trueage<-df1$age[df2$ID_unit]
	alk <- xtabs(~lenCat+age,data=df2%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
	rez1%>%mutate(n=nb)
}

#generate differente sample rate
rezall<-data.frame()
for(i in 1:30){
	rez0<-fct1(df1,i)
	rezall<-rbind(rez0,rezall)
}

#apply age length key to length data
#from FSA
prop<-1/0.01
rez1hist<-rezall%>%group_by(age,nbsamp=n)%>%summarise(n=n()*prop,type="sampalk")%>%ungroup()
rez2hist<-df1%>%group_by(age=trueage)%>%summarise(n=n()*prop,type="sampreal")%>%ungroup()
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")%>%ungroup()
#rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
#ggplot(rez1hist,aes(x=age,y=n,color=nbsamp))+geom_boxplot()+
ggplot(data=rez1hist,aes(x=age,y=n,group=age))+geom_boxplot()+
geom_point(data=rez1hist,aes(x=age,y=n,color=nbsamp,group=NULL),alpha=.2)+#geom_point(alpha=.2)+
	geom_line(data=rez2hist,aes(x=age,y=n,group=NULL),col="red")+
	geom_path(data=rez3hist,aes(x=age,y=n,group=NULL),col="green")






```

\newpage
# Session
```{r}
print(sessionInfo())
```
\newpage
# References




